---
title: "weather_report"
output:
  word_document: default
  html_document: default
---
## Loading libraries

```{r, message = FALSE}
library(tidyr)
library(dplyr)
library(stringr)
library(magrittr)
library(lubridate)
library(purrr)
```

## Loading data

```{r}
weather <- readRDS("~/R/rclass/weather.rds")
```

## Tidying data

Let's look at the data.

```{r}
str(weather)
```

We can see that the names of the columns are contained as rows in the 'Measure' column, and the columns that we have right now seem to be days of the month. So, our rows are our columns and our columns are our rows.

### Rotating the columns

```{r}
weather1 <- weather %>% select(-1) %>% 
  pivot_longer(cols = starts_with("X"), names_to = "day", values_to = "value") %>% 
  pivot_wider(names_from = "measure", values_from = "value")

weather1$day <- str_replace_all(weather1$day, "X", "")
```

```{r}
head(weather1)
tail(weather1)
str(weather1)
```

Now rows and columns are where they are supposed to be, and so the data it tidy, but still not clean. We have a bunch of NA's in our data, and the columns that should be numeric ('num' or 'int') are actually strings ('chr'). Also some of the values in the 'PrecipitationIn' column are 'T', even though it is not a logical variable, but a numeric one. Some of the cells in the 'Events' column are empty, we can assume that those denote the days when nothing happened.

We need to remove the missing values and values that are of wrong type, and change the 'empty' events to 'Nothing', so that it is clearer and easier to read and filter. Then we are going to change the data types of or columns.

### Removing missing values

```{r}
weather2 <- na.omit(weather1) 
weather2 %<>% filter(PrecipitationIn != "T") 
weather2$Events <- ifelse(str_length(weather2$Events) > 0, 
                          weather2$Events, 
                          "Nothing")

#str_replace_all would have been better, but it cannot use empty string as a pattern to replace

str(weather2)
```

## Fixing data types

### Fixing dates

First, we are going to fix the date. Right now it's in three separate columns, two of which are 'int' and one is 'chr'. We are going to clip them together and transform into a 'Date' type.

```{r}
weather2$Date = str_c(weather2$year, weather2$month, weather2$day, sep = ".") %>% ymd()
weather2 %<>% select(-c(1:3))
str(weather2)
```

### Fixing data types

We now will change the type of all columns except 'Events' and 'Date' to numeric. Events will be transformed into a factor, and Date is already good enough.

```{r}
weather3 = purrr::map_dfc(weather2[c(1:20, 22)], .f = as.numeric)
weather3$Date = weather2$Date
weather3$Events = as.factor(weather2$Events)
str(weather3)
```

That's better.

## Printing the result

```{r}
summary(weather3)
```

