---
title: "Weather Data 12.2014 - 12.2015"
author: "Eseniia Novokhatskaia" 
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Weather Report

This is an initial 'Weather' data set, which contains an information about different weather observations from December 2014 till December 2015. 

```{r}
library(tidyverse)
weather <- readRDS("/Users/apple/Desktop/weather.rds")
head(weather)
```

The problem with initial data set is you cannot really follow the logic of the data gathering and sorting, so it is needed to be fixed. 

First of all, we clearly see that 'X-' are days. We do not really need Xs, so we put numbers of days instead.

```{r}
weather$X <- NULL
names(weather)[4:34] <- '1':'31'
head(weather)
```

Then, we need to sort out the date of the observation. It is easier to unite day, month and year in one column and call it "date".

```{r}
weather <- weather %>%
  gather(day, data, '1': '31') %>%
  unite (Date, year, month, day)
head(weather)
```

Now we need to spread all the measures in different columns, so we have different dates as rows in our data set and different measures in columns.

```{r}
weather <- weather %>%
  spread(key = measure, value = data)
head(weather)
```

Weather data set already looks understandable, however, we can make it even better by sorting the data set by date and bringing all measure names to one style.

```{r}
weather <- weather %>%
  mutate(Date = as.Date(Date, format = "%Y_%m_%d")) %>%
  arrange(Date) %>%
  rename(
    Cloud.CoverOktas = CloudCover,
    Max.HumidityPercent = Max.Humidity, 
    Mean.HumidityPercent = Mean.Humidity, 
    Mean.Dew.PointF = MeanDew.PointF,
    Min.Dew.PointF = Min.DewpointF,
    Min.HumidityPercent = Min.Humidity,
    Wind.DirDegrees = WindDirDegrees,
  )
head(weather)
```



