---
title: "Tidy Weather"
author: "Baykov"
date: "02 12 2019"
output: html_document
---
# Connecting libraries and datasets
```{r message=FALSE}
library(dplyr)
library(tidyr)
weather <- readRDS("C:/Users/Baykov.f/Documents/R/weather.rds")
```

# Tidying data 

Cleaning dataset weather.

## 1) Exploring data

Understanding of structure. Let see what weather.rds is?
```{r}
class(weather)
dim(weather)
str(weather)
```

## 2) Now let's take a look at Data 

Here we see an example of default weather.rds
```{r}
head(weather)
```

## 3) Tidying data

#### First of all, lets get rid of column X
```{r}
weather1 <- select(weather, -X)
```

#### Now let's name columns "X1", "X2", etc to "1","2", etc
```{r}
names(weather1)[4:34] <- '1':'31'
```

#### We want to swap days and "meausre"
Also in procces we unite day, month, year in one single "d_m_y" date 
```{r}
weather2 <- gather(weather1, key = day, value = val, '1' : '31')
weather3 <- unite(weather2, col = d_m_y, day, month, year)
weather4 <- spread(weather3, key = measure, value = val)
weather4[c(1:6), c(1:6)]
```

#### Now let's deal with missing values

##### Before
```{r}
weather4[c(43:47), c(1:3,22)]
sum(is.na(weather4))
```
##### After
```{r}
weather4$Events[weather4$Events == ''] <- "Unknown"
weather4$PrecipitationIn[weather4$PrecipitationIn == 'T'] <- "0.00"
weather5 <- na.omit(weather4)
weather5[c(39:44), c(1:3,22)]
sum(is.na(weather5))
```
#### Change type of data in columns from "chr" to "num"
##### Before
```{r}
str(weather5)
```
##### After
```{r}
weather5[c(2,4:23)] <- lapply(weather5[c(2,4:23)], as.numeric)
str(weather5)
```
#### Finally ordering data by time
```{r}
weather5[c(1:10), c(1:5)]
weather6 <- weather5[order(as.Date(weather5$d_m_y, "%d_%m_%y")),] 
weather6[c(1:10), c(1:5)]
```

