---
title: "Weather"
author: "Stavrova Valeria"
date: "19 11 2019"
output:
  word_document: default
  pdf_document: default
---

# Loading data
```{r}
weather <-readRDS("C:/Users/Adm/Downloads/weather.rds")
```

# Explorying data
Let's have a look at the structure of the data.
```{r}
str(weather)
summary(weather)
```

# Tidying data
We see that values are stored in columns, not rows. We need to fix that. 
```{r}
library(tidyr)
library(dplyr)
newweather <-weather[,-1] %>% gather(key = day, value = val, X1 : X31)  %>% 
  pivot_wider(names_from = measure, values_from = val)  
```

Also we need to sort out the date of observation. I decided to keep days, months and years in different columns, because it may be convenient to select certain years/months/days as in the examples below. 

```{r}
library(data.table)
draftday<-data.table(newweather$day) 
day<- separate(draftday, V1, c("X", "day"),sep = 1)[,2]
finalweather<-cbind(day, newweather[,-3])[,c(2,3,1,4:25)]
```
We can also see that the columns that should be numeric are actually character. Precipitation column has some T values, but it may not be so, it's not a logical variable. We need to fix that too.

```{r}
finalweather1 <- finalweather %>% mutate_at(vars(-Events), as.numeric) %>% filter(PrecipitationIn != "T") 
```

Examples.
```{r}
finalweather1 %>% select(c(1:5)) %>% subset(year==2014)
finalweather1 %>% select(year, month, day, Events) %>% subset(month==12) %>% subset(day==1)
```

Then we deal with the missing values.

```{r}
finalweather2<-na.omit(finalweather1)
```

# Let's have a look at the result we've got
```{r}

summary(finalweather2)
```

# Visualising

```{r}
boxplot(finalweather2$Max.TemperatureF, horizontal = T)

boxplot(finalweather2$Max.Dew.PointF,finalweather1$MeanDew.PointF,finalweather1$Min.DewpointF)

hist(finalweather2$Min.Humidity)

boxplot(finalweather2$Mean.VisibilityMiles, horizontal = T)

hist(finalweather2$PrecipitationIn)
```

*The End*
